### 2023-10-22 assignment_2.qmd

---
title: "Assignment 2: Cross-sectional estimates"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
editor: visual
bibliography: [Econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
---

```{r}
#| label: setup
#| echo: false
#| output: false

library(tidyverse)
library(ggplot2)
library(magrittr)
library(dineq)
library(readxl)
```

```{r}
#| echo: false
#| output: false

# Adding data into dataset

# Data for population
pop_data <- read_csv("Raw_data/demo.gz")

# Data for GDP
gdp_data <- read_csv("Raw_data/nama.gz")
```

# Part A: Growth and Inequity

## 1. Model Estimation

## 2. Model Diagnostics

## 3. Visualization

# Part B: Exploring Other Determinants of Inequity

## 1. Data acquisition

## 2. Multiple Linear Regression Model

## 3. Model Interpretation